title: "courseproject1"
output: html_document
---

```{r setup, include=TRUE,echo=TRUE}
knitr::opts_chunk$set(echo = TRUE,results = "asis")
```

```{r reading data, echo=TRUE}
data <- read.csv("/Users/spandana/Desktop/activity.csv",header=TRUE)
```
## What is the mean total number of steps taken per day?

1.Calculate the total number of steps taken per day

```{r, echo=TRUE }
totalstepsperday <- aggregate(steps~date,data,sum,na.rm=TRUE)
```


2.Make a histogram of the total number of steps taken each day

```{r Figure1, echo=TRUE}
hist(totalstepsperday$steps,xlab="totalstepsperday",main="Histogram of totalstepsperday")
```
  
3. Calculate and report the mean and median of the total number of steps taken per day

```{r calculating mean and median, echo=TRUE}
meansteps <- mean(totalstepsperday$steps)
meansteps
mediansteps <- median(totalstepsperday$steps)
mediansteps
```
# What is the average daily activity pattern?

1. Make a time series plot of the 5-minute interval(x-axis) and the average number of steps taken, averaged across all days(y-axis)

```{r Figure2, echo=TRUE}
avgstepsperinterval <- aggregate(steps~interval,data=data,mean,na.rm=TRUE)
plot(steps~interval,data=avgstepsperinterval, type="l")
```

  
2. Which 5 minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r calculating the maximum of steps, echo=TRUE}
intervalwithsteps <- avgstepsperinterval[which.max(avgstepsperinterval$steps),]$interval
intervalwithsteps
```
# Inputting missing values 

1. Calculate and report the total number of missing values in the dataset

```{r calculating the total number of missing values, echo=TRUE}
totalmissingvalues <- sum(is.na(data$steps))
totalmissingvalues
```
  
2. Devise a strategy for filling in all of the missing values in the dataset 

Filling in the missing values in the dataset.For this purpose, first we will write a function below that will return the mean value for a particular interval

```{r fucntion to fill missing values,echo=TRUE }
getmeanvalueperinterval <- function(interval){
   avgstepsperinterval[avgstepsperinterval$interval==interval,]$steps
}
```  
  
3. Create a new dataset that is equal to the original dataset but with the missing data filled in

```{r creating a new dataset,echo=TRUE}
newdata<-data
for(i in 1:nrow(newdata)){
if(is.na(newdata[i,]$steps)){
newdata[i,]$steps <- getmeanvalueperinterval(newdata[i,]$interval)
}
}

head(newdata)
```

4. Make a histogram of the total number of steps taken each day and calculate and report the mean and median total number of steps taken per day. Do these values differ from the first part of the assignment? What is the impact of inputting missing data on the estimates of the total daily number of steps?

```{r Figure3,echo=TRUE}
totalstepsperday_newdata <- aggregate(steps ~ date, newdata, sum)
hist(totalstepsperday_newdata$steps,xlab="totalstepsperday",main="Histogram of total steps per day using newdata")
```

 
```{r calculating mean and median steps using newdata,echo=TRUE}
meansteps_newdata <- mean(totalstepsperday_newdata$steps)
meansteps_newdata
mediansteps_newdata <- median(totalstepsperday_newdata$steps)
mediansteps_newdata
```
  
# Are there differences in activity patterns between weekdays and weekends?
  
1.Create a new factor variable in the dataset with two levels -"weekday" and "weekend" indicating whether a given data is a weekday or weekend dat

```{r creating weekday and weekend variables,echo=TRUE}
newdata$date <- as.Date(strptime(newdata$date, format="%Y-%m-%d"))
newdata$day <- weekdays(newdata$date)
for(i in 1:nrow(newdata)){
  if(newdata[i,]$day %in% c("Saturday","Sunday")){
    newdata[i,]$day <- "weekend"
  }
  else{
    newdata[i,]$day<-"weekday"
  }
}
stepsperday<-aggregate(newdata$steps ~ newdata$interval + newdata$day, newdata, mean)
```
  
2.Make a panel plot containing a time series plot of the 5 minute interval and the avg number of steps taken,averaged across all weekdat days or weekend days. 

```{r Figure4, echo=TRUE}
names(stepsperday) <- c("interval","day","steps")
library(lattice)
xyplot(steps~interval | day, stepsperday, type="l", layout = c(1,2), xlab="Interval", ylab="Number of steps")
```


>>>>>>> first commit
